---
title: "003Markdown_TwoFactorial"
output: html_document
date: "2024-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instal Packages
```{r instal packages}

install.packages("forcats")
install.packages("DHARMa")
library(ggplot2)# plotting
library(dplyr)# data management and summary statistics
library(ggpubr)# plotting
library(openxlsx)# import and export Excel files
library(forcats)# factor repordering
library(DHARMa) # model diagnostics
library(emmeans) # posthoc tests
library(multcomp) # cld
library(multcompView) #cld
library(glmmTMB) # mixed model
library(car) # anova for glmmTMB
library(lme4) # lmer and glmer mixed model
library(lmerTest) # test lmer
library(conflicted)
conflicts_prefer(lmerTest::lmer)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

# Import Data
```{r Import Data}
df = read.xlsx("rawData/data_YSM.xlsx")

str(df)
head(df)
tail(df)
```
# Data transformation

```{r}
df <- df %>%
  mutate(across(c(geno, N), ~ as.factor(.x)),
         N = fct_relevel(N, "low", "med", "high"),
         block = as.factor(rep))



str(df)
summary(df)
!is.na(df)
class(df)

```
 # 1st. Plot
 
```{r}
ggplot(data = df, aes(x= N, y= yield, col = geno)) +
  stat_boxplot(geom = "errorbar", width = 0.6) +
  geom_boxplot(outlier.shape=NA, 
               width = 0.6) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.4, jitter.height = 0, 
                                            dodge.width=0.6), shape=1, size=1) 

```
# Fit Model

```{r}
model = lm(yield ~ geno * N, data = df)


summary(model)

```
Yield of geno4 at high N: Intercept + genog4 + Nhigh + genog4:Nhigh


# Check Model Assumptions
## Heteroscedasticity
-- Dharma: residual diagnosics

```{r}
simulationOutput <- simulateResiduals(fittedModel = model, plot = F)
plot(simulationOutput)
```
# within group Residual variability
```{r}
plotResiduals(simulationOutput, form = df$geno)
```
# Test of significance and posthoc tests

```{r}
anova(model)
```
```{r}
emmeans(model, ~geno*N)

```
# Groupwise comparison
```{r}
contrast(emmeans(model, ~N|geno), method="pairwise", infer=c(T,T)) 
```

```{r}
emmeans(model, ~N|geno)
```
# vergleich Standard vs Kontrolle
```{r}
contrast(emmeans(model, ~N|geno), method="trt.vs.ctrl") # erstes Level = Standard
contrast(emmeans(model, ~N|geno), method="trt.vs.ctrl", ref = 3) #drittes level= standard

# Ein Test weniger, kann statistisch sinnvll sein/ helfen Signifikanz zu erreichen
```
```{r}
cld(emmeans(model,  ~N|geno ), method= "pairwise", adjust="sidak", Letters=letters)
```
# Presentation of results

```{r}
CIs=cld(emmeans(model, ~N|geno), method="pairwise", adjust="sidak", Letters=letters)
CIs$.group =gsub(" ", "", CIs$.group, fixed = TRUE)
CIs



ggplot(df, aes(y=yield, x=N, col=N)) +
  geom_point(shape=1, size=1.5)+
  geom_point(data=CIs, aes(y=emmean), 
             shape=16,  size=2, 
             position = position_nudge(x = 0.4))+
  geom_errorbar(data=CIs, aes(y=emmean, ymin=lower.CL, ymax=upper.CL), 
                width=0.2, position = position_nudge(x = 0.4))+
  geom_text(data=CIs, aes(y = 110, label =.group), col=1)+
 scale_color_manual(values=c(rgb(240,215,35, max = 255), 
                             rgb(190,210,35, max = 255), 
                              rgb(35,80,150, max = 255)                              ),
                     limits = c("low","med", "high"))+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "none")+
  ylab("Dry matter yield (dt/ha)")+
  xlab("Nitrogen fertilisation")
```

# Analysis of Pests in relation to N, Genotype

```{r}

ggplot(df, aes(y = pests, x = N, col = geno)) +
  stat_boxplot(geom ="errorbar", width = 0.6)+
  geom_boxplot(outlier.shape=NA, width = 0.6) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, jitter.height = 0, 
                                            dodge.width=0.6), shape=1, size=1)
```



```{r}
model_pests = lm(pests ~ geno*N, data= df)


simulationOutput <- simulateResiduals(fittedModel = model_pests, plot = F)
plot(simulationOutput)
```
```{r}
plotResiduals(simulationOutput, form = df$geno)
```
```{r}
plotResiduals(simulationOutput, form = df$N)
```

# Test of Significance

```{r}
anova(model_pests)
```

```{r}
contrast(emmeans(model_pests, ~N|geno), 
         #method = "pairwise"
         method="trt.vs.ctrl"
         )
```

```{r}
cld(emmeans(model_pests, ~N|geno), 
    #method="pairwise", 
    method="trt.vs.ctrl",
    adjust="sidak", Letters=letters)
```
```{r}
CIs=cld(emmeans(model_pests, ~N|geno), method="pairwise", adjust="sidak", Letters=letters)
CIs$.group =gsub(" ", "", CIs$.group, fixed = TRUE)
CIs



ggplot(df, aes(y=pests, x=N)) +
  #geom_boxplot(col = geno) +
  geom_point(shape=1, size=1.5)+
  geom_point(data=CIs, aes(y=emmean), 
             shape=16,  size=2, 
             position = position_nudge(x = 0.4))+
  geom_errorbar(data=CIs, aes(y=emmean, ymin=lower.CL, ymax=upper.CL), 
                width=0.2, position = position_nudge(x = 0.4))+
  geom_text(data=CIs, aes(y = 110, label =.group), col=1)+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "none")+
  ylab("pests")+
  xlab("Nitrogen fertilisation")
```
# Linear Model as fixed Effect

```{r}
df$block=as.factor(df$rep)
ggplot(df, aes(y=yield, x=N, col=N, shape=block)) +
  geom_point(size=1.5)+
  scale_color_manual(values=c(rgb(240,215,35, max = 255), 
                             rgb(190,210,35, max = 255), 
                              rgb(35,80,150, max = 255)),
                     limits = c("low","med", "high"))+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "bottom")+
  ylab("Dry matter yield (dt/ha)")+
  xlab("Nitrogen fertilisation")
```
```{r}
model2 <- lm(yield ~ geno * N + block, data = df)
summary(model2)
```

```{r with Block}
CI2s=cld(emmeans(model2, ~N|geno), method="pairwise", adjust="sidak", Letters=letters)
CI2s$.group =gsub(" ", "", CI2s$.group, fixed = TRUE)
CI2s

ggplot(df, aes(y=yield, x=N, col=N)) +
  geom_point(shape=1, size=1.5)+
  geom_point(data=CI2s, aes(y=emmean), 
             shape=16,  size=2, 
             position = position_nudge(x = 0.4))+
  geom_errorbar(data=CI2s, aes(y=emmean, ymin=lower.CL, ymax=upper.CL), 
                width=0.2, position = position_nudge(x = 0.4))+
  geom_text(data=CI2s, aes(y = 110, label =.group), col=1)+
 scale_color_manual(values=c(rgb(240,215,35, max = 255), 
                             rgb(190,210,35, max = 255), 
                              rgb(35,80,150, max = 255)                              ),
                     limits = c("low","med", "high"))+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "none")+
  ylab("Dry matter yield (dt/ha)")+
  xlab("Nitrogen fertilisation")
```

# Compare Model with and w/out Block Effekt
```{r compare with and without block}
CI.comp=rbind(CIs, CI2s)
CI.comp$model=rep(c("model", "model2"), each=12)

ggplot(data=CI.comp, aes(y=emmean, x=N, col=model, group=model))+
  geom_point(position=position_dodge(width=0.4))+
  geom_errorbar(aes(y=emmean, ymin=lower.CL, ymax=upper.CL), 
                width=0.4, 
                position=position_dodge(width=0.4))+
  facet_grid(~geno)+
  theme(legend.position="bottom")
```

# Linear mixed effect model with block as random effect (glmmTMB)

```{r}
model3 <- glmmTMB(yield ~ geno * N + (1|block), data = df, REML=T)
summary(model3)

Anova(model3)

CI3s=cld(emmeans(model3, ~N|geno), method="pairwise", adjust="sidak", Letters=letters)
CI3s$.group =gsub(" ", "", CI3s$.group, fixed = TRUE)
CI3s
```
# Compare CIs of various models



```{r}
CI.comp=rbind(CIs, CI2s, CI3s, CI4s)
CI.comp$model=factor(rep(c("lm crd", "lm rcbd", "glmmTMB",  "lmer"), each=12), levels =c("lm crd", "lm rcbd", "glmmTMB",  "lmer"))
```


# gmm for pest.a in dependance of genotype and N
```{r}
model_glm = glmmTMB(pests.A ~ geno * N + (1|block), data = df, family="poisson")

summary(model_glm)
```


# have a look at data

# Test model

```{r}
simulationOutput <- simulateResiduals(fittedModel = model_glm, plot = F)
testDispersion(simulationOutput)
```


# Test for disperion

```{r}
plot(simulationOutput)
```

family = negative binominial (nbinom2)

```{r}
model_nbinom2 = glmmTMB(pests.A ~ geno * N + (1|block), data = df, family="nbinom2")

summary(model_nbinom2)
```
```{r}
simulationOutput <- simulateResiduals(fittedModel = model_nbinom2, plot = F)
testDispersion(simulationOutput)
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = model, plot = F)
plot(simulationOutput)
```
```{r}
plotResiduals(simulationOutput, form = df$N)
```


# plot model

