---
title: "004Mrkdwn_GLM"
output: html_document
date: "2024-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}
library(ggplot2)# plotting
library(dplyr)# data management and summary statistics
library(ggpubr)# plotting
library(openxlsx)# import and export Excel files
library(forcats)# factor repordering
library(DHARMa) # model diagnostics
library(emmeans) # posthoc tests
library(multcomp) # cld
library(multcompView) #cld
library(glmmTMB) # mixed model
library(car) # anova for glmmTMB
library(lme4) # lmer and glmer mixed model
library(lmerTest) # test lmer
library(conflicted)
conflicts_prefer(lmerTest::lmer)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

```{r Load data}
df3 = read.xlsx("data_YSM.xlsx")
str(df3)
```

Aus irgendeinem Grund konnte data_YSM.xlsx nicht aus dem Unterordner gelesen werden (?!!!), darum Daten direkt in Ordner mit Code geladen

```{r Data transformation}
df <- df3 %>%
  mutate(across(c(geno, N), ~ as.factor(.x)),
         N=fct_relevel(N, "low", "med", "high"),
         block=as.factor(rep))
str(df)
```

# GLM for countdata

```{r define Model}
model2=glm(pests ~ geno * N + block, data = df, family="poisson")
summary(model2)
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = model2, plot = F)
testDispersion(simulationOutput)
```

```{r}
plot(simulationOutput)
```

```{r}
plotResiduals(simulationOutput, form = df$geno)
```

```{r}
plotResiduals(simulationOutput, form = df$N)
```

```{r}
plotResiduals(simulationOutput, form = df$block)
```

```{r}
emmeans(model2, ~N*geno)
```

```{r}
emmeans(model2, ~N*geno, type = "response")
```

```{r}
contrast(emmeans(model2, ~N|geno
                 , type="response"), method="trt.vs.ctrl", infer=c(T,T))
```

```{r}
contrast(emmeans(model2, ~N|geno), method="trt.vs.ctrl", infer=c(T,T))
```

```{r Confidence Intervals}
CI2s=cld(emmeans(model2, ~N|geno, type="response"), method="pairwise", adjust="sidak", Letters=letters)
CI2s$.group =gsub(" ", "", CI2s$.group, fixed = TRUE)
CI2s
```

```{r}
ggplot(df, aes(y=pests, x=N, col=N)) +
  stat_boxplot(geom ="errorbar", width = 0.6)+
  geom_boxplot(outlier.shape=NA, width = 0.6) +
  geom_jitter(position=position_jitterdodge(jitter.width = 0.1, jitter.height = 0, 
                                            dodge.width=0.6), shape=1, size=1)+
  geom_point(data=CI2s, aes(y=rate), 
             shape=16,  size=2, 
             position = position_nudge(x = 0.4))+
  geom_errorbar(data=CI2s, aes(y=rate, ymin=asymp.LCL, ymax=asymp.UCL), 
                width=0.2, position = position_nudge(x = 0.4))+
  geom_text(data=CI2s, aes(y = 60, label =.group), col=1)+
 scale_color_manual(values=c(rgb(240,215,35, max = 255), 
                             rgb(190,210,35, max = 255), 
                              rgb(35,80,150, max = 255)                              ),
                     limits = c("low","med", "high"))+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "none")+
  ylab("Number of pests per trap and 48h")+
  xlab("Nitrogen fertilisation")
```

```{r}
ggplot(df, aes(y=pests, x=N, col=N)) +
  geom_point(shape=1, size=1.5)+
   geom_point(data=CI2s, aes(y=rate), 
             shape=16,  size=2, 
             position = position_nudge(x = 0.4))+
  geom_errorbar(data=CI2s, aes(y=rate, ymin=asymp.LCL, ymax=asymp.UCL), 
                width=0.2, position = position_nudge(x = 0.4))+
  geom_text(data=CI2s, aes(y = 60, label =.group), col=1)+
 scale_color_manual(values=c(rgb(240,215,35, max = 255), 
                             rgb(190,210,35, max = 255), 
                              rgb(35,80,150, max = 255)                              ),
                     limits = c("low","med", "high"))+
  facet_grid(~geno, scales="free", labeller = labeller(geno=label_both))+
  theme_bw() +
  theme(legend.position = "none")+
  ylab("Number of pests per trap and 48h")+
  xlab("Nitrogen fertilisation") +
  scale_y_sqrt(breaks=c(0,3,10,20,40,60))
```

# Connect Git

## install if needed (do this exactly once):

install.packages("usethis")

library(usethis) 
use_git_config(user.name = "KostaKostaLemon", user.email = "kobra-bradke@web.de")

#Configure the default name for an initial branch
usethis::git_default_branch_configure()

